\section{Discussion}
\label{sec:discussion}
The results support a restrained conclusion. UL-TransXNet is a competitive compact ultrasound classifier with favorable multi-dataset in-domain behavior, but it should not be described as uniformly superior on every dataset and every metric. On TN5000, ConvNeXt-Tiny remains slightly stronger in the repeated-seed table; on AUL, Swin-T slightly exceeds UL-TransXNet in macro F1 and raw accuracy despite lower AUC and balanced accuracy. The proposed model becomes clearly advantageous in the multi-dataset view: it is substantially stronger on BUSI, has higher AUC and balanced accuracy on AUL, and has the best average performance across the three public benchmarks under dataset-specific training.

The central story is therefore not that three generic modules are stacked onto TransXNet and happen to improve an average table. A more defensible interpretation is that UL-TransXNet organizes compact local--global feature modeling around an ROI workflow for ultrasound lesions, then tests the resulting classifier under four linked views: controlled oracle ROI classification, validation-audited model-family selection, automatic-ROI substitution, and mobile prototype inference. This framing matches the evidence more closely than a broad claim of a new universal backbone.

The TN5000 ablations clarify the role of the architecture components. The structure-level progression shows that \texttt{TransXNet-GG} is the strongest TN5000 variant, while the \texttt{TransXNet-GGG-MCA} configuration used by UL-TransXNet is the best three-dataset choice in the validation audit and final result tables. The module-level ablation shows that differential attention is the clearest contributor to TN5000 thresholded metrics. MUDD has a smaller and less uniform effect. MCA does not uniformly improve TN5000 AUC, but it improves TN5000 F1 and accuracy and produces large BUSI gains. This means that MCA should be framed as an operating-point and multi-dataset behavior module rather than as the sole reason for the model family performance.

This interpretation is important in medical image analysis. Ultrasound datasets differ substantially in organ anatomy, acquisition quality, lesion size, and class prevalence. A small AUC difference on one thyroid benchmark is less informative than the pattern across several public benchmarks and operating points. The evidence suggests that the proposed architecture is useful because it gives a favorable balance across dataset-specific evaluations, not because every module monotonically improves every metric.

The automatic ROI experiments address a separate practical concern. The main classifier is trained under an annotation-derived ROI protocol, but the TN5000, BUSI, and AUL closed-loop evaluations show that detector-predicted crops can replace oracle crops with a moderate performance loss and remain clearly stronger than full-image classification where that comparison is evaluated. This result should be interpreted as workflow evidence rather than detector novelty. In particular, AUL localization is weaker than BUSI localization, which explains part of the remaining gap between automatic and oracle ROI classification.

The Android results remain a practical feasibility check. Many ultrasound classification papers stop at workstation-side accuracy. The mobile experiment shows that a TN5000-calibrated prototype can run in a low-latency mode while preserving strong classification quality, and that a more expensive ensemble mode is available when latency is less critical. These measurements do not constitute clinical deployment validation or prove superiority over all mobile baselines, but they show that UL-TransXNet is compatible with edge-side inference constraints.


\subsection*{Limitations}
This study has several limitations. First, the evaluation is still based on public retrospective datasets, and label definitions, class proportions, and preprocessing conditions are not fully uniform across datasets. BUSI and AUL are dataset-specific in-domain evaluations rather than thyroid-to-breast or thyroid-to-liver external generalization tests. Second, the BUSI/AUL protocol uses five training folds evaluated on a fixed held-out test set, so the reported standard deviations reflect training-fold variation rather than an outer cross-validation estimate. Third, the work focuses on binary benign--malignant classification rather than detection, segmentation, or multi-class clinical diagnosis. Fourth, the automatic ROI experiment uses a practical one-class detector only to form the classifier input; it does not establish a new lesion-detection method, and the AUL results show that localization quality can still limit closed-loop performance. Fifth, the mobile experiment demonstrates prototype feasibility but not a full clinical workflow under realistic operator interaction and acquisition variability. Finally, external robustness remains an open problem because the evaluation is still limited to public retrospective datasets rather than prospective multi-center cohorts. Future studies should therefore focus on broader external validation, stronger detector-classifier integration, calibration under distribution shift, and clinically grounded prospective evaluation.
