\section{Conclusion}
\label{sec:conclusion}
This paper presents UL-TransXNet, a compact ROI framework for benign--malignant ultrasound lesion classification. The evidence supports a restrained but coherent conclusion. UL-TransXNet is competitive on TN5000, clearly stronger on BUSI and AUL, and strongest on average across the three public datasets under dataset-specific training and the unified reporting protocol. The validation audit explains why the final model is selected for three-dataset behavior rather than for the single best TN5000 row. TN5000 ablations show that the architecture gains are not explained by a single module: the GG structure is the strongest thyroid-only variant, differential attention improves thresholded behavior, and MCA mainly improves multi-dataset operating behavior. Automatic ROI experiments further show that detector-predicted crops can connect the classifier to a practical localization front end and remain stronger than full-image classification, including a TN5000 automatic-ROI sanity check for the primary benchmark. Android experiments show that the model can support prototype speed--accuracy trade-offs on mobile hardware. At the same time, broad external generalization and clinical deployment remain unresolved because all reported datasets are public retrospective benchmarks. Future studies should therefore focus on larger external validation, stronger automatic ROI generation, more rigorous calibration under distribution shift, and deployment evaluation under realistic clinical workflows.
