\documentclass[review,12pt]{elsarticle}

\usepackage{amsmath,amssymb}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{adjustbox}
\usepackage{subcaption}
\usepackage{siunitx}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{float}
\usepackage{url}

\journal{Pattern Recognition}
\graphicspath{{figures/}}
\hypersetup{colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue}
\setlength{\emergencystretch}{2em}

\begin{document}

\begin{frontmatter}

\title{UL-TransXNet: A Lightweight Architecture for Multi-Dataset Ultrasound Lesion Classification}

\author[inst1]{Zihan Li}
\author[inst1]{Cong Yang\texorpdfstring{\corref{cor1}}{}}
\ead{cong.yang@suda.edu.cn}
\cortext[cor1]{Corresponding author.}

\affiliation[inst1]{organization={School of Future Science and Engineering, Soochow University},
            city={Suzhou},
            postcode={215222},
            state={Jiangsu},
            country={China}}

\begin{abstract}
Ultrasound lesion classification is challenging because speckle noise, weak boundaries, low contrast, and acquisition-dependent variation obscure lesion cues. We present UL-TransXNet, a compact ROI-based TransXNet family architecture for benign--malignant ultrasound lesion classification. Lesion annotations are used to construct expanded square ROI crops, after which a lightweight hierarchical backbone models local texture and broader contextual relationships. We evaluate the model on three public datasets: TN5000 for thyroid nodules, BUSI for breast lesions, and AUL for liver lesions. Under the unified ROI protocol, UL-TransXNet obtains mean test AUC/balanced-accuracy pairs of 0.9483/0.8726 on TN5000, 0.8998/0.7961 on BUSI, and 0.8767/0.8200 on AUL. The result does not imply uniform superiority: ConvNeXt-Tiny remains slightly stronger on TN5000 at the selected operating point. Instead, the main finding is cross-dataset competitiveness, with the strongest average performance across the three datasets and clear gains on BUSI and AUL. TN5000 ablations show that the TransXNet family refinements improve operating-point behavior, while MCA contributes mainly to thresholded decision quality and cross-dataset robustness rather than to a uniform AUC gain. An auxiliary automatic-ROI experiment further shows that detector-predicted crops improve full-image classification on BUSI and AUL, with a remaining gap to oracle ROI performance. Android experiments quantify deployment trade-offs, where horizontal-flip test-time augmentation improves AUC from 0.9402 to 0.9466 at 226.2 ms per image. These results support UL-TransXNet as a competitive lightweight ultrasound classifier, while broader domain generalization remains open.
\end{abstract}

\begin{highlights}
\item UL-TransXNet classifies ultrasound lesions from ROI crops.
\item Three public datasets support cross-organ evaluation.
\item Ablations isolate MCA, MUDD, and differential attention.
\item Automatic ROI experiments connect localization and classification.
\item Android experiments quantify accuracy--latency trade-offs.
\end{highlights}

\begin{keyword}
Ultrasound image classification \sep Thyroid nodule \sep Breast ultrasound \sep Liver ultrasound \sep Lightweight neural network \sep Mobile deployment
\end{keyword}

\end{frontmatter}

\input{sections/01_introduction}
\input{sections/02_related_work}
\input{sections/03_method}
\input{sections/04_experimental_setup}
\input{sections/05_results}
\input{sections/06_discussion}
\input{sections/07_conclusion}
\input{sections/08_appendix}

\section*{CRediT authorship contribution statement}
Zihan Li: Conceptualization, methodology, software, experiments, and writing---original draft. \\
Cong Yang: Supervision, conceptualization, writing---review and editing, resources, and project administration.

\section*{Ethics statement}
This study is a secondary analysis of publicly available, de-identified ultrasound datasets. No new patient data were collected by the authors. Ethical approval and consent procedures are governed by the original dataset publications or data providers; therefore, no additional institutional review board approval was required for this retrospective computational study.

\section*{Funding}
This research did not receive any specific grant from funding agencies in the public, commercial, or not-for-profit sectors.

\section*{Declaration of competing interest}
The authors declare that they have no known competing financial interests or personal relationships that could have appeared to influence the work reported in this paper.

\section*{Data availability}
The datasets analyzed in this study are publicly available from their original sources, which are cited in the manuscript. The source code, paper-generation scripts, Android demonstration source, and compact result summaries are available at \url{https://github.com/Afr1ste/ul-transxnet}. Dataset files, trained checkpoints, ONNX binaries, and Android application packages are not redistributed in the repository; they are archived with the project materials and are available from the corresponding author upon reasonable request, subject to the applicable redistribution terms of the original datasets.

\bibliographystyle{elsarticle-num}
\bibliography{refs}

\end{document}

